Ages of Stars and Planets Observed by Kepler

### 2024-04-21 paper/ms.tex
update ms

%\title{Gyrochrone and Lithium Ages for Stars and Planets Observed by Kepler}
\title{Ages of Planets and Stars Observed by Kepler}

@@

\subsection{High Resolution Spectra}
\label{subsec:lithiumsel}
A final piece of our analysis involves an age assessment based on the
equivalent width of the \ion{Li}{1} 6708\,\AA\ doublet.  For this
component of the work, we rely on spectra acquired using the High
Resolution Echelle Spectrometer (HIRES; \citealt{vogt_hires_1994}) on
the Keck I 10m telescope.  These spectra were collected between 2009
and 2024, primarily through the California Kepler Survey
\citep{2017AJ....154..107P,2017AJ....154..108J,2017AJ....154..109F}.
We supplemented the existing archive with $\approx$10~hours of new
observations for 22 stars between Fall 2022 and Spring 2024.
These stars were chosen to ensure that confirmed planets with
rotational evidence for ages below 1\,Gyr received spectra, since this
is the age range where lithium is most likely to yield useful age
constraints.

Lithium equivalent widths and abundances were already analyzed by
\citet{2018ApJ...855..115B} for about three quarters of the spectra in
our sample.
% stat from _estimate_new_lithium_timefrac.py
However, since new spectra have been acquired since the
\citeauthor{2018ApJ...855..115B} analysis, and since our approach and
selection function are different, we opt to perform our own analysis
of the reduced HIRES spectra.

%FIXME FIXME TODO LEFT OFF HERE
%FIXME FIXME TODO LEFT OFF HERE
%FIXME FIXME TODO LEFT OFF HERE
These HIRES spectra are available for \nkoiswithhires\ of the
\nkoisnofp\ non-false-positive KOIs.
In the broader stellar sample, they are available for a much smaller 
fraction of the stars observed by Kepler, because they were generally only
acquired due to the presence of a transiting planet candidate.
Given this selection function,
we leverage these spectra in our analysis as a way to scrutinize the
ages of just the planet sample.

@@

{\it Candidate pulsators and close-in binaries (bit
10)}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g. RR~Lyrae, Cepheids) and candidate
close-in binaries.  Visual inspection hints that this flag also
includes many bona fide young stars, and so although we propagate it 
to our set of quality flags, we generally ignore it.

@@

We queried the internal CPS archive to collect all blaze-corrected
HIRES spectra that were available for the
\nplhoststarwgyroagewithgrazingandhighruwe\ stars for which
rotation-based ages can be derived.  This query yielded at least one
spectrum for \nlithiumstars\ and \nlithiumplanets\ planets.  These
spectra were acquired between 6 September 2009 and 16 May 2024.  For
stars with multiple spectra available at varying ranges of
signal-to-noise (S/N), we analyzed only the spectrum with the highest
number of counts.

We measured the lithium equivalent widths following a procedure from
previous work \citep{Bouma_2021}.  The stars of interest for this work
are FGK stars, and so the continuum in the vicinity of the \ion{Li}{1}
6708\,\AA\ doublet is well-defined.  We therefore numerically
integrate a single best-fit Gaussian over a local window, and
estimated uncertainties on the line width through a Monte Carlo
procedure that bootstraps against the local scatter in the continuum.
This approach does not correct for the neighboring \ion{Fe}{1}
6707.44\,\AA\ blend.

To evaluate the accuracy and precision of our equivalent measurement
technique, after applying an initial iteration on the
\nplhoststarwgyroagewithgrazingandhighruwe\ aforementioned stars,
we compared our lithium equivalent widths
with those reported by \citet{2018ApJ...855..115B}.
For the \nbergeroverlap\ stars in both samples, we found {\bf X, Y,
Z.}

>>>>

%\title{Gyrochrone and Lithium Ages for Stars and Planets Observed by Kepler}
%Ages of the Kepler Stars and Planets
\title{Ages of Stars and Planets Observed by Kepler}

@@

\subsection{High Resolution Spectra}
\label{subsec:lithiumsel}

The final piece of our analysis involves assessing ages based on the
equivalent width of the \ion{Li}{1} 6708\,\AA\ doublet.  Specifically,
we analyzed spectra acquired with the High Resolution Echelle
Spectrometer (HIRES; \citealt{vogt_hires_1994}) on the Keck I 10m
telescope.  These spectra were primarily collected through the
California Kepler Survey
\citep{2017AJ....154..107P,2017AJ....154..108J,2017AJ....154..109F}.
We supplemented the existing archive with $\approx$10~hours of new
observations for 22 stars between Fall 2022 and Spring 2024.  These
stars were chosen to ensure that confirmed planets with rotational
evidence for ages below 1\,Gyr had spectra, since this is the age
range in which lithium is most likely to yield useful age constraints.

% stat from _estimate_new_lithium_timefrac.py
Lithium equivalent widths and abundances for the Kepler Objects of
Interest were already analyzed by \citet{2018ApJ...855..115B} for
about three quarters of the spectra in our sample.  However, since new
spectra have been acquired since the \citeauthor{2018ApJ...855..115B}
analysis, and since our approach and selection function are different,
we opt to perform our own analysis of the reduced HIRES spectra.

We queried the internal CPS archive to collect all blaze-corrected
HIRES spectra that were available for all non-false positive Kepler
Objects of Interest with a MES$>$10.  This query yielded at least one
spectrum for \nlithiumstars\ stars and \nlithiumplanets\ planets.  For
stars with multiple spectra available at varying ranges of
signal-to-noise (S/N), we analyzed only the spectrum with the highest
number of counts.  The resulting spectra were acquired between 6
September 2009 and 16 May 2024, and cumulatively comprise
\nhireshours\,hours of open-shutter time.

We measured the lithium equivalent widths following a procedure
adapted from previous work \citep{Bouma_2021}.  Our stars of interest
are FGK stars, and so the continuum in the vicinity of the \ion{Li}{1}
6708\,\AA\ doublet is well-defined.  We Doppler corrected the spectra
to a common reference wavelength by cross-correlating against a high
S/N template for Kepler-1698, chosen because $\gamma$$\approx$9\kms,
$T_{\rm eff}$$\approx$5000\,K and $v\sin i$$\approx$5\,\kms, which
puts it in the middle of our sample's temperature range and gives it
mild line-broadening.  We then trimmed the Doppler-corrected spectra
to a local window centered on the lithium line, using a window width
of 15\,\AA (we considered 10\,\AA\ and 20\,\AA; the results were
broadly consistent).  We then continuum-normalized by fitting a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines.  We then numerically integrated the resulting
spectrum using a one-component Gaussian with free amplitude, width,
and mean, and estimated uncertainties on the line width through a
Monte Carlo procedure that bootstraped against the local scatter in
the continuum.

%    %FIXME FIXME TODO LEFT OFF HERE
%    %FIXME FIXME TODO LEFT OFF HERE
%    %FIXME FIXME TODO LEFT OFF HERE
Our EW measurement approach does not correct for the neighboring
\ion{Fe}{1} 6707.44\,\AA\ blend.
To evaluate the accuracy and precision of our equivalent measurement
technique, after applying an initial iteration on the
\nplhoststarwgyroagewithgrazingandhighruwe\ aforementioned stars,
we compared our lithium equivalent widths
with those reported by \citet{2018ApJ...855..115B}.
For the \nbergeroverlap\ stars in both samples, we found {\bf X, Y,
Z.}

%    These HIRES spectra are available for \nkoiswithhires\ of the
%    \nkoisnofp\ non-false-positive KOIs.
%    In the broader stellar sample, they are available for a much smaller 
%    fraction of the stars observed by Kepler, because they were generally only
%    acquired due to the presence of a transiting planet candidate.
%    Given this selection function,
%    we leverage these spectra in our analysis as a way to scrutinize the
%    ages of just the planet sample.

@@

{\it Candidate pulsators and close-in binaries (bit
10)}---\citeauthor{Santos_2021} included in a flag for candidate
``classical pulsators'' (e.g. RR~Lyrae, Cepheids) and candidate
close-in binaries.  Visual inspection of the rotation periods as a
function of temperature shows that although this flag selects many
bona fide objects in these classes, it also selects the overwhelming
majority of $\lesssim$5\,day bona fide young planets.  While we
propagate this flag into our set of quality flags, we therefore
generally ignore it.

@@
